---
title: 'Unsupervised Learning Mini-Project'
author: "Shreya Sunkara"
date: "10/29/2019"
output: html_document
---

```{r}
input.file <- "https://bioboot.github.io/bimm143_F19/class-material/WisconsinCancer.csv"

#wisc.df <- read.csv("WisconsinCancer.csv")
wisc.df <- read.csv(input.file)
head(wisc.df)
```

```{r}
nrow(wisc.df)
```

How many benign (not cancerous) and and malignant (cancerous) samples do we have in the dataset?

```{r}
table(wisc.df$diagnosis)
```

```{r}
# Convert the features of the data: wisc.data
wisc.data <- as.matrix(wisc.df[,3:32])

# Set the row names of wisc.data
row.names(wisc.data) <- wisc.df$id

head(wisc.data)
```

Store the diagnosis for reference in the future as a separate vector

```{r}
diagnosis <- wisc.df$diagnosis
```

How many observations are in this dataset?
```{r}
nrow(wisc.df)
```

How many of the observations have a malignant diagnosis?
```{r}
table(wisc.df$diagnosis)
```

How many variables/features in the data are suffixed with _mean?
```{r}
colnames(wisc.df)
```

```{r}
grep("_mean", colnames(wisc.df), value = TRUE)
```

I can use `length()` to count how many matches we have.

```{r}
length( grep("_mean", colnames(wisc.df)) )
```

# Performing PCA

```{r}
round( colMeans(wisc.data), 3)
```

```{r}
round( apply(wisc.data, 2, sd), 3)
```

```{r}
# Perform PCA on wisc.data 
wisc.pr <- prcomp( wisc.data, scale= TRUE )
summary(wisc.pr)
```

```{r}
plot(wisc.pr)
```

Lets make a plot of PC1 vs PC2 

```{r}
plot(wisc.pr$x[,1],  wisc.pr$x[,2])
```

Color by cancer/non-cancer...
```{r}
plot(wisc.pr$x[,1],  wisc.pr$x[,2], col=diagnosis,
     xlab="PC1", ylab="PC2")
```




```{r}
x <- summary(wisc.pr)
x$importance[,"PC1"]
#x$importance[,1]
```

How many principal components (PCs) are required to describe at least 70% of the original variance in the data?

```{r}
x$importance[3,]
```

How many principal components (PCs) are required to describe at least 90% of the original variance in the data?

```{r}
which(x$importance[3,] > 0.9 )[1]
```

#Factoextra package 
```{r}
## ggplot based graph
#install.packages("factoextra")
library(factoextra)
fviz_eig(wisc.pr, addlabels = TRUE)
```

## Communicating PCA results

For the first principal component, what is the component of the loading vector (i.e. wisc.pr$rotation[,1]) for the feature concave.points_mean?


```{r}
wisc.pr$rotation["concave.points_mean",1]
```

What is the minimum number of principal components required to explain 80% of the variance of the data?

```{r}
x <- summary(wisc.pr)
which(x$importance["Cumulative Proportion",] > 0.8)[1]
```

## Hierarchical clustering

```{r}
wisc.hclust <- hclust( dist( scale(wisc.data)) )
plot(wisc.hclust)
```

```{r}
wisc.pr.hclust <- hclust( dist(wisc.pr$x[,1:7]),
                          method="ward.D2")
```


```{r}
plot(wisc.pr.hclust)
```

```{r}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
```



```{r}
table(grps, diagnosis)
```

#Prediction

```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

#Plot new samples on PC1 vs PC2 plot

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=diagnosis)
points(npc[,1], npc[,2], col="blue", pch=15, cex=3)
text(npc[,1], npc[,2], labels = c(1,2), col="white")
```

