---
title: "Class14"
author: "Shreya Sunkara"
date: "11/12/2019"
output: html_document
---

##DESeq2 Setup

We will use the DESeq2 package from Bioconductor. To install this package, we ran in the console. (Note: eval=FALSE)

```{r eval=FALSE}
#install.packages("BiocManager")
#BiocManager::install()
#BiocManager::install("DESeq2")
```
##Input Dataset

Input countData and colData (metadata)

```{r}
counts <- read.csv("airway_scaledcounts.csv",stringsAsFactors = FALSE)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```
Have a peak...
```{r}
head(counts)
```
```{r}
head(metadata)
```

How many genes do we have in this dataset?
```{r}
nrow(counts)
```
How many experiments do we have in this dataset?
```{r}
ncol(counts)
```
```{r}
nrow(metadata)
```

```{r}
control <- metadata[metadata[,"dex"]=="control",]
```

Now, calculate mean values across control columns of **countdata**
```{r}
control.mean <- rowSums(counts[,control$id])/length(control$id)
names(control.mean) <- counts$engene
```
Now, do the same thing for the treated columns of **countdata**
```{r}
treated.mean <- rowSums(counts[,treated$id])/length(control$id)
names(treated.mean) <- counts$engene
```
```{r}
meancounts <- data.frame(control.mean, treated.mean)
plot(meancounts)
```

Next, we want to show these data on a log scale
```{r}
plot(meancounts$control.mean,meancounts$treated.mean,log="xy")
```
Positive fold change would mean it's upregulated by drug; negative fold change would mean it's downregulated by drug.
```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```
Now, we want to weed out the NaN and Inf values
```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```
To see how many genes are up and down regulated
```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```
```{r}
sum(up.ind)
```
```{r}
head(mycounts[up.ind,])
```
# DESeq2 Analysis
Now, to properly analyze the data. DESeq2 is an R package
```{r}
library(DESeq2)
```
```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```

Get results
```{r}
res <- results(dds)
res
```
```{r}
plot(res$log2FoldChange,-log(res$padj))
abline(v=c(2,2),col="gray",lty=2)

```
```{r}
write.csv(res,file="expression_results.csv")
```







